---
pagetitle: "Reading GeoJSON files"
lang: "en-GB"
output:
  html_document:
    highlight: textmate
    css: ../../styles.css
    includes:
      in_header: ../../head_includes.html
      before_body: ../../header.html
      after_body: ../../footer.html
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<!-- Add a title -->
<h1>Reading GeoJSON files</h1>

<!-- Provide a short summary -->
<p>Read a GeoJSON file, transform the coordinates and plot.</p>

<!-- List the R packages needed and a link to their page on cran.r-project.org -->
<div class = "row">
<div class = "col-md-6">
<div class="titlebar">R packages</div>
<ul>
  <li><a href="https://cran.r-project.org/web/packages/sf/index.html" target="_blank">sf</a></li>
</ul>
</div>

<!-- Provide a link to the data source -->
<div class = "col-md-6">
<div class="titlebar">Data</div>
<ul>
  <li><a href="https://geoportal.statistics.gov.uk/datasets/countries-december-2020-uk-buc" target="_blank">Countries of the UK</a> (ONS Open Geography Portal)</li>
</ul>
</div>
</div>

<!-- Add each stage of the workflow -->
<div class="titlebar">Workflow</div>

<p>First we'll load the <a href="https://cran.r-project.org/web/packages/sf/index.html" target="_blank">sf</a> package.</p>
```{r}
library(sf)
```

<p>Then we'll read the GeoJSON file by passing its URL to the <a href="https://r-spatial.github.io/sf/reference/st_read.html" target="_blank"><code>st_read()</code></a> function.</p>
```{r}
sf <- st_read("https://opendata.arcgis.com/datasets/bb7104d3a9c04937be57e408288282dc_0.geojson")
```

<p>The GeoJSON file is read into R and information about the spatial vector boundaries is printed in the console. We can see that the data are projected in WGS84 (EPSG:4326). This is what we'd expect for a <a href="https://datatracker.ietf.org/doc/html/rfc7946#page-12" target="_blank">GeoJSON file</a>. However, we want the data to be projected in British National Grid (EPSG:27700). To reproject the data we can use the <a href="https://r-spatial.github.io/sf/reference/st_transform.html" target="_blank"><code>st_transform()</code></a> function and supply the <a href="https://epsg.org/home.html" target="_blank">Coordinate Reference System</a> to the `crs` argument.</p>
```{r}
sf_bng <- st_transform(sf, crs = 27700)
```

<p>To check that the data is projected in the right CRS we can inspect the CRS units. WGS84 uses decimal degrees (lat,long) whereas British National Grid uses metres.</p>
```{r}
st_crs(sf)$units_gdal
st_crs(sf_bng)$units_gdal
```

<p>Finally, we'll plot the data to see if it covers the area of interest.</p>
```{r}
plot(st_geometry(sf), col = "lightgrey")
```

<div class="titlebar">Notes</div>
<p>When a coordinate reference system (CRS) is missing or wrongly set, the <a href="https://r-spatial.github.io/sf/reference/st_crs.html" target="_blank"><code>st_set_crs()</code></a> function can be used.</p>
<br /><br />