---
pagetitle: "Reading Excel files"
lang: "en-GB"
output:
  html_document:
    highlight: textmate
    css: ../../styles.css
    includes:
      in_header: ../../head_includes.html
      before_body: ../../header.html
      after_body: ../../footer.html
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<!-- Add a title -->
<h1>Reading Excel files</h1>

<!-- Provide a short summary -->
<p>Load an Excel file, select a specific worksheet and skip unnecessary rows.</p>

<!-- List the R packages needed and a link to their page on cran.r-project.org -->
<div class = "row">
<div class = "col-md-6">
<div class="titlebar">R packages</div>
<ul>
  <li><a href="https://cran.r-project.org/web/packages/readxl/index.html" target="_blank">readxl</a></li>
</ul>
</div>

<!-- Provide a link to the data source -->
<div class = "col-md-6">
<div class="titlebar">Data</div>
<ul>
  <li><a href="https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatesforukenglandandwalesscotlandandnorthernireland" target="_blank">Median age of population for UK local authorities</a> (ONS)</li>
</ul>
</div>
</div>

<!-- Add each stage of the workflow -->
<div class="titlebar">Workflow</div>

<p>First load the <a href="https://cran.r-project.org/web/packages/readxl/index.html" target="_blank">readxl</a> package.</p>
```{r}
library(readxl)
```

<p>Then create an object with the path to the Excel file.</p>
```{r}
path <- "../../data/ukpopestimatesmid2020on2021geography.xls"
```

<p>Next we'll list all of the worksheets in the spreadsheet using the <a href="https://readxl.tidyverse.org/reference/excel_sheets.html" target="_blank"><code>excel_sheets()</code></a> function.</p>
```{r}
excel_sheets(path)
```

<p>We are interested in the worksheet called "MYE 6" which contains the median age of local authority populations in the UK. So let's use the <a href="https://readxl.tidyverse.org/reference/read_excel.html" target="_blank"><code>read_excel()</code></a> function and specify "MYE 6" in the `sheet` argument to load the worksheet.</p>
```{r}
df <- read_excel(path = path, sheet = "MYE 6")
```

<p>Inspecting the first few rows of the data we notice that there is a lot of metadata.</p>
```{r}
df
```

<p>So we'll read the worksheet again but use the `skip` argument to remove the first 7 rows.</p>
```{r}
df <- read_xls(path = path, sheet = "MYE 6", skip = 7)
```

<p>To check that the data are tidy we can print the first few rows with the `head()` function.</p>
```{r}
head(df)
```

<div class="titlebar">Notes</div>
<p>We have read the Excel file locally but you may want to download it directly from a URL. To do this we can use the <a href="https://httr.r-lib.org/reference/GET.html" target="_blank"><code>GET()</code></a> function from the <a href="https://cran.r-project.org/web/packages/httr/index.html" target="_blank">httr</a> package.</p>
```{r, eval = FALSE}
library(httr)
url <- "https://www.ons.gov.uk/file?uri=%2fpeoplepopulationandcommunity%2fpopulationandmigration%2fpopulationestimates%2fdatasets%2fpopulationestimatesforukenglandandwalesscotlandandnorthernireland%2fmid2020/ukpopestimatesmid2020on2021geography.xls"
GET(url, write_disk(temp <- tempfile(fileext = ".xls")))
df <- read_xls(temp, sheet = "MYE 6", skip = 7)
```
<br /><br />